---
title: "LAGOS Spatial Analysis"
author: "Matthew Ross"
date: "9/11/2019"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
library(ggthemes)
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)

```{r data-read}
# #Lagos download script
LAGOSNE::lagosne_get(dest_folder = LAGOSNE:::lagos_path())
# LAGOSNE compiled to C:\Users\13074\AppData\Local/LAGOSNE/LAGOSNE//data_1.087.3.qs

#Load in lagos (Per instructions from  LAGOS. see ?lagosne_load for details)
lagos <- lagosne_load()

#Grab the lake centroid info
lake_centers <- lagos$locus

#load not working.
#load('lake_centers.RData')
#load('C:/Users/13074/Documents/ESS580/LAGOS/data/lake_centers.Rdata')

```



### Convert to spatial data
```{r}
#Look at the column names
#names(lake_centers)

#Look at the structure
#str(lake_centers)

#View the full dataset, (1:100 is first row, 100 columns)
#View(lake_centers %>% slice(1:100))

# function st_as_sf is simple table and turn it into a simple feature, which is a spatial object. When using this, longitudes are "x", lattitudes are "y"

#st_transform Transform or convert coordinates of simple feature

# crs is a EPSG projection code (google "WGS 84 EPSG to get projection systems), 4326 is WGS 84 

spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)

#Subset for plotting
subset_spatial <- spatial_lakes %>%
  slice(1:100) 

#???
#subset_baser <- spatial_lakes[1:100,]

#Dynamic mapviewer
mapview(subset_spatial)

```


### Subset to only Minnesota

```{r}
states <- us_states()

#Plot all the states to check if they loaded
mapview(states)
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)

#st_transform changes the projection

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,]

#Plotting the first 1000 lakes
minnesota_lakes %>%
  arrange(-lake_area_ha) %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')
```



# In-Class work


## 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r}
#subsetting Iowa and Illinois
iowa_illinois <- states %>%
  filter(name %in% c('Iowa', 'Illinois')) %>%
  st_transform(2163)

# map
mapview(iowa_illinois)

```



## 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa
combined? How does this compare to Minnesota?

*Combined, Iowa and Illinois have 16466 lakes, while Minnesota has an additional 12572 lakes for a total of 29038 lakes by itself.* 
```{r}

# subsetting LAGOS data for Iowa and Illinois
ia_il_lakes <- spatial_lakes[iowa_illinois,]

# Counting obeservations in the two dfs
nrow(ia_il_lakes)
nrow(minnesota_lakes)

nrow(minnesota_lakes)-nrow(ia_il_lakes)

```


## 3) What is the distribution of lake size in Iowa vs. Minnesota?

- Here I want to see a histogram plot with lake size on x-axis and frequency on 
y axis (check out geom_histogram)

*Minnesota has a much larger distribution of lakes at all sizes than Iowa, with the largest lakes being nearly five times the largest size of Iowa.* 
```{r}

# Using st_join to join the spatial object and sf dataframes. 
# subsetting states
ia_mn <- states %>%
  filter(name %in% c('Iowa', 'Minnesota')) %>%
  st_transform(2163)

# joining the two 
ia_mn_lakes <- st_join(spatial_lakes, ia_mn) %>% 
  filter(!is.na(name))

# Plot the two states
ggplot(ia_mn_lakes,aes(lake_area_ha, fill = name, color = name)) +
  geom_histogram(alpha = 0.8) +
  facet_wrap(~name) +
  scale_x_log10() +
  theme_base() +
  xlab("Lake Area (ha)") +
  ylab("Count")

```


## 4) Make an interactive plot of lakes in Iowa and Illinois and color them
by lake area in hectares

*Plot shows the largest 500 lakes according to size (ha) in Iowa and Illinois.* 
```{r}

#Plotting the top 500 largest lakes by hectares
ia_il_lakes %>%
  arrange(-lake_area_ha) %>%
    slice(1:500) %>%
  mapview(.,zcol = 'lake_area_ha', layer.name = "Lakes (ha)")


```


## 5) What other data sources might we use to understand how reservoirs and 
natural lakes vary in size in these three states? 

*It appears that the USGS's Water Resources Mission Area had spatial data related to reservoir and lake size, as well as the NWIS Mapper and NSDI Water Node. The EPA's LAKECAT seems to have related data too.*


